### 2016-11-28 results.tex
safjas;fja;lfjas;ld
formatting changes
section on sub-optimal network pruning

\begin{figure}[!h]
\centering
\includegraphics[width=0.49\linewidth]{png/cos-small.pdf}
\includegraphics[width=0.49\linewidth]{png/cos-big.pdf}
\caption{Degradation in squared error after pruning a two-layer network trained to compute the cosine function (\textbf{Left Network:} 2 layers, 10 neurons each, 1 output, logistic sigmoid activation, starting test accuracy: 0.9999993, \textbf{Right Network:} 2 layers, 50 neurons each, 1 output, logistic sigmoid activation, starting test accuracy: 0.9999996)}
\label{fig:cosine-double-layer}
\end{figure}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-single-pass-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-single-pass-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a single-layer network using The Single Overall Ranking algorithm (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-single-ranking-single-layer}
\end{figure}

@@

\subsubsection{Iterative Re-Ranking Algorithm}

\begin{figure}[!hb]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-iterative-rerank-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-iterative-rerank-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a single-layer network the Iterative Re-ranking algorithm (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-re-ranking-single-layer}
\end{figure}

In Figure \ref{fig:mnist-re-ranking-single-layer} we present our results using Algorithm \ref{algo2} (The Iterative Re-Ranking Algorithm) in which all remaining neurons are re-ranked after each successive neuron is switched off. We compute the same brute force rankings and Taylor series approximations of error deltas over the remaining active neurons in the network after each pruning decision. This is intended to account for the effects of cancelling interactions between neurons.

There are 2 key observations here. Using the Brute Force ranking criteria, almost 60\% of the neurons in the network can be pruned away without any major loss in performance. The other noteworthy observation here is that the 2nd order Taylor Series approximation of the error performs consistently better than its 1st order version.

@@

\begin{figure}[!hb]
\centering
\includegraphics[width=\linewidth]{png/mnist-acc99-gt-gain.pdf}
\caption{Error surface of the network output in log space (left) and in real space (right) with respect to each candidate neuron chosen for removal using the brute force criterion; (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-gt-single-layer}
\end{figure}

\begin{figure}[!hb]
\centering
\includegraphics[width=\linewidth]{png/mnist-acc99-g1-gain.pdf}
\caption{Error surface of the network output in log space (left) and in real space (right) with respect to each candidate neuron chosen for removal using the 1st order Taylor Series error approximation criterion; (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-gt-single-layer}
\end{figure}

\begin{figure}[!hb]
\centering
\includegraphics[width=\linewidth]{png/mnist-acc99-g2-gain.pdf}
\caption{Error surface of the network output in log space (left) and in real space (right) with respect to each candidate neuron chosen for removal using the 2nd order Taylor Series error approximation criterion; (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-gt-single-layer}
\end{figure}

@@

We graph the effect of boosting the neuron's output to demonstrate that for certain neurons in the network, even doubling, tripling, or quadrupling the scalar output of the neuron has no effect on the overall error of the network, indicating the remarkable degree to which the network has learned to ignore the value of certain parameters. In other cases, we can get a sense of the sensitivity of the network's output to the value of a given neuron when the curve rises steeply after the red 1.0 line. This indicates that the learned value of the parameters emanating from a given neuron are relatively important, and this is why we should ideally see sharper upticks in the curves for the later-removed neurons in the network, that is, when the neurons crucial to the learning representation start to be picked off.

%Remember that lower is better in terms of the height of the curve and minimal horizontal change between the vertical red line at 1.0 (neuron \textit{on}, $\alpha = 1.0$) and 0.0 (neuron \textit{off}, $\alpha = 0.0$) is indicative of a good candidate neuron to prune, i.e. there will be minimal effect on the network output when the neuron is removed.

Some very interesting observations can be made in each of these graphs.

\textbf{Brute Force Criterion}

Notice how low to the floor and flat most of these curves are. It's only until the 90th removed neuron that we see a higher curve with a more convex shape (clearly a more influential piece of the network). This again validates the fact that a good majority of the neurons in a network do not contribute to the overall performance.

\textbf{1st Order Approximation Criterion}

It can be seen that most choices seem to have flat or negatively sloped curves, indicating that the first order approximation seems to be pretty good, but examining the brute force choices shows they could be better.

\textbf{2nd Order Approximation Criterion}

This method looks much more similar to the Brute Force method choices, though clearly not as good (they're more spread out). Notice the difference in convexity between the 2nd and 1st order method choices. It's clear that the first order method is fitting a line and the 2nd order method is fitting a parabola in their approximation.

\subsection{Pruning A 2-Layer Network}
The network architecture in this case consisted of 2 layers, 50 neurons per layer, 10 outputs, logistic sigmoid activations, and a starting test accuracy of 1.000.
\subsubsection{Single Overall Ranking Algorithm}

\begin{figure}[!hb]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-deep-single-pass-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-deep-single-pass-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a 2-layer network using the Single Overall Ranking algorithm; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-single-ranking-double-layer}
\end{figure}

@@

Unsurprisingly, a 2-layer network is harder to prune because a single overall ranking will never capture the interdependencies between neurons in different layers. It makes sense that this is much worse than the performance on the 1-layer network, even if this method is already known to be bad, and we'd likely never use it in practice.

\subsubsection{Iterative Re-Ranking Algorithm}

\begin{figure}[!hb]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-deep-iterative-rerank-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-deep-iterative-rerank-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a 2-layer network using the Iterative Re-ranking algorithm; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-re-ranking-double-layer}
\end{figure}

@@

\begin{figure}[!hb]
\centering
\includegraphics[width=\linewidth]{png/mnist-deep-gt-gain.pdf}
\caption{Error surface of the network output in log space (left) and in real space (right) with respect to each candidate neuron chosen for removal using the brute force criterion; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-gt-double-layer}
\end{figure}

\begin{figure}[!hb]
\centering
\includegraphics[width=\linewidth]{png/mnist-deep-g1-gain.pdf}
\caption{Error surface of the network output in log space (left) and in real space (right) with respect to each candidate neuron chosen for removal using the 1st order Taylor Series error approximation criterion; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-g1-double-layer}
\end{figure}

\begin{figure}[!hb]
\centering
\includegraphics[width=\linewidth]{png/mnist-deep-g2-gain.pdf}
\caption{Error surface of the network output in log space (left) and in real space (right) with respect to each candidate neuron chosen for removal using the 2nd order Taylor Series error approximation criterion; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-g2-double-layer}
\end{figure}

As seen in the case of a single layered network, these graphs are a visualization the error surface of the network output with respect to the neurons chosen for removal using each algorithm, represented in intervals of 10 neurons.

\textbf{Brute Force Criterion}

It is clear why these neurons got chosen, their graphs clearly show little change when neuron is removed, are mostly near the floor, and show convex behaviour of error surface, which argues for the rationalization of using 2nd order methods to estimate difference in error when they are turned off.

\textbf{1st Order Approximation Criterion}

@@

\textbf{2nd Order Approximation Criterion}

Clearly these neurons are not overtly poor candidates for removal (error doesn't change much between 1.0 \& zero-crossing left-hand-side), but could be better (as described above in the Brute Force Criterion discussion).

@@

\begin{figure}[!hb]
\centering
\includegraphics[width=0.80\linewidth]{png/diamond.png}
\includegraphics[width=0.80\linewidth]{png/rshape.png}
\caption{Degradation in squared error after pruning a 2-layer network using the Single Pass Algorithm (left) and the Iterative Re-ranking algorithm(right) on toy "diamond" shape dataset (above) and toy "random shape" dataset (below); (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.992(diamond); 0.986(random shape)}
\label{fig:diamond}
\end{figure}

As can be seen from the experiments on MNIST, even though the 2nd-order approximation criterion is consistently better than 1st-order, it's performance is not nearly as good as brute force based ranking, especially beyond the first layer. What is interesting to note is that from some other experiments conducted on toy datasets (predicting whether a given point would lie inside a given shape on the Cartesian plane), the performance of the 2nd-order method was found to be exceptionally good and produced results very close to the brute force method. The 1st-order method, as expected, performed poorly here as well. Some of these results are illustrated in Figure \ref{fig:diamond}. This huge variation in results from MNIST might be attributed to the fact that these toy datasets had only 2 output classes (as opposed to 10 classes in MNIST), but it certainly warrants further investigation.

%\begin{figure}[!hb]
%\centering
%\includegraphics[width=0.80\linewidth]{png/rshape.png}
%\caption{Degradation in squared error after pruning a 2-layer network using the Single Pass Algorithm (left) and the Iterative Re-ranking algorithm(right) on toy "random shape" dataset; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.986)}
%\label{fig:rshape}
%\end{figure}

%It was also observed that the brute force method primarily prunes neurons from the outer layer initially, which suggests that it might be a good idea to look into pruning neurons based on the layer they belong to.

%From some other experiments conducted on smaller toy datasets, it was noted that the pruning performance is worse if the models are not trained to a high accuracy, which is quite expected since this simply means that in these cases all the neurons have significant contribution to the network's performance and hence there is still scope for a better accuracy given more neurons and/or increased training times. However, pruning in these cases typically improved the MSE for the first 5 to 10 \% of neurons removed which can again be attributed to the theory put forward by \cite{mozer1989using}.

>>>>

\begin{figure}[!ht]
\centering
\includegraphics[width=0.49\linewidth]{png/cos-small.pdf}
\includegraphics[width=0.49\linewidth]{png/cos-big.pdf}
\caption{Degradation in squared error after pruning a two-layer network trained to compute the cosine function (\textbf{Left Network:} 2 layers, 10 neurons each, 1 output, logistic sigmoid activation, starting test accuracy: 0.9999993, \textbf{Right Network:} 2 layers, 50 neurons each, 1 output, logistic sigmoid activation, starting test accuracy: 0.9999996)}
\label{fig:cosine-double-layer}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-single-pass-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-single-pass-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a single-layer network using The Single Overall Ranking algorithm (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-single-ranking-single-layer}
\end{figure}

@@

\subsubsection{Iterative Re-Ranking Algorithm}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-iterative-rerank-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-acc99-iterative-rerank-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a single-layer network the iterative re-ranking algorithm (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-re-ranking-single-layer}
\end{figure}

In Figure \ref{fig:mnist-re-ranking-single-layer} we present our results using Algorithm \ref{algo2} (The iterative re-ranking Algorithm) in which all remaining neurons are re-ranked after each successive neuron is switched off. We compute the same brute force rankings and Taylor series approximations of error deltas over the remaining active neurons in the network after each pruning decision. This is intended to account for the effects of cancelling interactions between neurons.

There are 2 key observations here. Using the brute force ranking criteria, almost 60\% of the neurons in the network can be pruned away without any major loss in performance. The other noteworthy observation here is that the 2nd order Taylor Series approximation of the error performs consistently better than its 1st order version, in most situations, though Figure \ref{fig:mnist-test-single-digit-2} is a poignant counter-example.

@@

@@

We graph the effect of boosting the neuron's output to demonstrate that for certain neurons in the network, even doubling, tripling, or quadrupling the scalar output of the neuron has no effect on the overall error of the network, indicating the remarkable degree to which the network has learned to ignore the value of certain parameters. In other cases, we can get a sense of the sensitivity of the network's output to the value of a given neuron when the curve rises steeply after the red 1.0 line. This indicates that the learned value of the parameters emanating from a given neuron are relatively important, and this is why we should ideally see sharper upticks in the curves for the later-removed neurons in the network, that is, when the neurons crucial to the learning representation start to be picked off. Some very interesting observations can be made in each of these graphs.

Remember that lower is better in terms of the height of the curve and minimal (or negative) horizontal change between the vertical red line at 1.0 (neuron \textit{on}, $\alpha = 1.0$) and 0.0 (neuron \textit{off}, $\alpha = 0.0$) is indicative of a good candidate neuron to prune, i.e. there will be minimal effect on the network output when the neuron is removed.

\subsubsection{Visualization of brute force Pruning Decisions}
\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-acc99-gt-gain.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the brute force criterion; (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-single-layer-gt}
\end{figure}

In Figure \ref{fig:mnist-gt-single-layer}, we notice how low to the floor and flat most of the curves are. It's only until the 90th removed neuron that we see a higher curve with a more convex shape (clearly a more sensitive, influential piece of the network).

\subsubsection{Visualization of 1st Order Approximation Pruning Decisions}
\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-acc99-g1-gain.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the 1st order Taylor Series error approximation criterion; (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-single-layer-g1}
\end{figure}

It can be seen in Figure \ref{fig:mnist-single-layer-g1} that most choices seem to have flat or negatively sloped curves, indicating that the first order approximation seems to be pretty good, but examining the brute force choices shows they could be better.

\subsubsection{Visualization of 2nd Order Approximation Pruning Decisions}
\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-acc99-g2-gain.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the 2nd order Taylor Series error approximation criterion; (\textbf{Network:} 1 layer, 100 neurons, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.998)}
\label{fig:mnist-single-layer-g2}
\end{figure}

The method in Figure \ref{fig:mnist-single-layer-g2} looks similar to the brute force method choices, though clearly not as good (they're more spread out). Notice the difference in convexity between the 2nd and 1st order method choices. It's clear that the first order method is fitting a line and the 2nd order method is fitting a parabola in their approximation.

\subsection{Pruning A 2-Layer Network}
The network architecture in this case consisted of 2 layers, 50 neurons per layer, 10 outputs, logistic sigmoid activations, and a starting test accuracy of 1.000.

\subsubsection{Single Overall Ranking Algorithm}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-deep-single-pass-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-deep-single-pass-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a 2-layer network using the Single Overall Ranking algorithm; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-single-ranking-double-layer}
\end{figure}

@@

Unsurprisingly, a 2-layer network is harder to prune because a single overall ranking will never capture the interdependencies between neurons in different layers. It makes sense that this is worse than the performance on the 1-layer network, even if this method is already known to be bad, and we'd likely never use it in practice.

\subsubsection{iterative re-ranking Algorithm}
\begin{figure}[!ht]
\centering
\includegraphics[width=0.49\linewidth]{png/mnist-deep-iterative-rerank-method.pdf}
\includegraphics[width=0.49\linewidth]{png/mnist-deep-iterative-rerank-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a 2-layer network using the iterative re-ranking algorithm; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-re-ranking-double-layer}
\end{figure}

@@

\subsubsection{Visualization of Error Surface \& Pruning Decisions}
As seen in the case of a single layered network, these graphs are a visualization the error surface of the network output with respect to the neurons chosen for removal using each algorithm, represented in intervals of 10 neurons.

\subsubsection{Visualization of brute force Pruning Decisions}
\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-deep-gt-gain.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the brute force criterion; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-gt-double-layer}
\end{figure}

In Figure \ref{fig:mnist-gt-double-layer}, it is clear why these neurons got chosen, their graphs clearly show little change when neuron is removed, are mostly near the floor, and show convex behaviour of error surface, which argues for the rationalization of using 2nd order methods to estimate difference in error when they are turned off.

\subsubsection{Visualization of 1st Order Approximation Pruning Decisions}
\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-deep-g1-gain.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the 1st order Taylor Series error approximation criterion; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-g1-double-layer}
\end{figure}

@@

\subsubsection{Visualization of 2nd Order Approximation Pruning Decisions}
\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-deep-g2-gain.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the 2nd order Taylor Series error approximation criterion; (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 1.000)}
\label{fig:mnist-g2-double-layer}
\end{figure}

Clearly these neurons are not overtly poor candidates for removal (error doesn't change much between 1.0 \& zero-crossing left-hand-side), but could be better (as described above in the brute force Criterion discussion).

\subsection{Investigation of Pruning Performance With Imperfect Starting Conditions}

In our experiments thus far we have tacitly assumed that we start with a network which has learned an ``optimal'' representation of the training objective, i.e. it has been trained to the point where we accept its performance on the test set. Here we explore what happens when we prune with a sub-optimal starting network.

If the assumptions of this paper regarding the nature of neural network learning are correct, we expect that two processes are essentially at work during back-propagation training. First, we expect that the neurons which directly participate in the fundamental learning representation (even if redundantly) work together to reduce error on the training data. Second, we expect that neurons which do not directly participate in the learning representation work to cancel each other's negative influence. Furthermore, we expect that these two groups are essentially distinct, as evinced by the fact that multiple neurons can often be removed as a group with little to no effect on the network output. Some non-trivial portion of the training time, then, is spent doing work which has nothing intrinsically to do with the learning representation and essentially functions as noise cancellation.

If this is the case, when we attempt to prune a network which has not fully canceled the noisy influence of extraneous or redundant units, we might expect to see the error actually \textit{improve} after removing a few bad apples. This is in fact what we observe, as demonstrated in the following experiments.

For each experiment in this section we trained with the full MNIST training set (\cite{lecun-mnisthandwrittendigit-2010}), uncompressed and without any data normalization. We trained three different networks to learn to distinguish a single handwritten digit from the rest of the data. The network architectures were each composed of 784 inputs, 1 hidden layer with 100 neurons, and 2 soft-max outputs; one to say yes, and the other to say no. These networks were trained to distinguish the digits 0, 1, and 2, and their respective starting accuracies were a sub-optimal 0.9757, 0.9881, and 0.9513. Finally, we only consider the iterative re-ranking algorithm, as the single overall ranking algorithm is clearly nonviable.

\subsubsection{MNIST Single Digit Classification: Digit 0}

Figure \ref{fig:mnist-test-single-digit-0} shows the degradation in squared error after removing neurons from a network trained to distinguish the digit 0. What we observe is that the first and second order methods both fail in different ways, though clearly the second order method makes better decisions overall. The first order method explodes spectacularly in the first few iterations. The brute force method, in stark contrast, actually \textit{improves} in the first few iterations, and remains essentially flat until around the 60\% mark, at which point it begins to gradually increase and meet the other curves.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\linewidth]{png/mnist-test-single-digit-0.pdf}
\caption{Degradation in squared error after pruning a single-layer network trained to do a one-versus-all classification of the digit 0 using the iterative re-ranking algorithm}
\label{fig:mnist-test-single-digit-0}
\end{figure}

The behavior of the brute force method here demonstrates that the network was essentially working to cancel the effect of a few bad neurons when the training convergence criteria were met, i.e. the network was no longer able to make progress on the training set. After removing these neurons during pruning, the output improved. We can investigate this by looking at the error surface with respect to the neurons chosen for removal by each method in turn. Below in Figure \ref{fig:mnist-test-single-digit-0-gt} is the graph of the brute force method.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-0-gt.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the brute force iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-0-gt}
\end{figure}

Figure \ref{fig:mnist-test-single-digit-0-gt} shows an interesting phenomenon, which we will see in later experiments as well. The high blue curve corresponding to neuron 0 is negatively sloped in the beginning and clearly after removing this neuron, the output will improve. The rest of the curves, in correspondence with the squared error degradation curve above, are mostly flat and tightly layered together, indicating that they are good neurons to remove.

In Figure \ref{fig:mnist-test-single-digit-0-g1} below, we observe a stark contrast to this. The curves corresponding to neurons 0 and 10 are mostly flat, and fairly lower than the rest, though clearly a mistake was made early on and the rest of the curves are clearly bad choices. In all of these cases however, we see that the curves are easily approximated with a straight line and so the first order method may have been fairly accurate in its predictions, even though it still made poor decisions.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-0-g1.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the first-order iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-0-g1}
\end{figure}

Figure \ref{fig:mnist-test-single-digit-0-g1} is an example of how things can go south once a few bad mistakes are made at the outset. Figure \ref{fig:mnist-test-single-digit-0-g2} shows a much better set of choices made by the second order method, though clearly not as good as the brute force method. The log-space plots make it a bit easier to see the difference between the brute force and second order methods in Figures \ref{fig:mnist-test-single-digit-0-gt} and \ref{fig:mnist-test-single-digit-0-g2}, respectively.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-0-g2.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the second-order iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-0-g2}
\end{figure}

\subsubsection{MNIST Single Digit Classification: Digit 1}

Examining Figure \ref{fig:mnist-test-single-digit-1}, we see a much starker example of the previous phenomenon, in which the brute force method continues to \textit{improve} the performance of the network after removing 80\% of the neurons in the network. The first and second order methods fail early and proceed in fits and starts (clearly demonstrating evidence of interrelated groups of noise-canceling neurons), and never fully recover. It should be noted that it would be impossible to see curves like this if neural networks evenly distributed the learning representation evenly or equitably over their hidden units.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\linewidth]{png/mnist-test-single-digit-1.pdf}
\caption{Degradation in squared error after pruning a single-layer network trained to do a one-versus-all classification of the digit 1 using the iterative re-ranking algorithm}
\label{fig:mnist-test-single-digit-1}
\end{figure}

One of the most striking things about the blue curve in Figure \ref{fig:mnist-test-single-digit-1} is the fact that the network never drops below its starting error until it crosses the 80\% mark, indicating that only 20\% of the neurons in this network are actually essential to the learning the training objective. In this sense, we can only wonder how much of the training time was spent winnowing the error out of the remaining 80\% of the network.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-1-gt.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the brute force iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-1-gt}
\end{figure}

In Figures \ref{fig:mnist-test-single-digit-1-gt}, \ref{fig:mnist-test-single-digit-1-g1} and \ref{fig:mnist-test-single-digit-1-g2} we can examine the choices made by the respective methods. The brute force method serves as our example of a near-optimal pruning regimen, and the rest are first and second order approximations of this. Small differences, clearly, can lead to large effects on the network output as shown in Figure \ref{fig:mnist-test-single-digit-1}.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-1-g1.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the first-order iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-1-g1}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-1-g2.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the second-order iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-1-g2}
\end{figure}

\subsubsection{MNIST Single Digit Classification: Digit 2}

Figure \ref{fig:mnist-test-single-digit-2} is an interesting case because it shatters our confidence in the reliability of the second order method to make good pruning decisions, and further demonstrates the phenomenon of how much the error can improve if the right neurons are removed after training gets stuck. In this case, though still a poor performance overall, the first order method vastly outperforms the second order method.

Figure \ref{fig:mnist-test-single-digit-2-gt} shows us a clear example of the first element to remove having a negative error slope, and improving the output as a result. The rest of the pruning decisions are reasonable. Comparing with the blue curve in Figure \ref{fig:mnist-test-single-digit-2}, we see the correspondence between the first pruning decision improving the output, and the remaining pruning decisions keeping the output fairly flat. Clearly, however, there isn't much room to get worse given our starting point with a sub-optimal network, and we see that the ending sum of squared errors is not much higher than the starting point. At the same time, we can still see the contrast in performance if we make optimal pruning decisions, and most of the neurons in this network were clearly doing nothing.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\linewidth]{png/mnist-test-single-digit-2.pdf}
\caption{Degradation in squared error after pruning a single-layer network trained to do a one-versus-all classification of the digit 2 using the iterative re-ranking algorithm}
\label{fig:mnist-test-single-digit-2}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-2-gt.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the brute force iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-2-gt}
\end{figure}

In Figure \ref{fig:mnist-test-single-digit-2-g1}, we see a mixed bag in which the decisions are clearly sub-optimal, though much better than Figure \ref{fig:mnist-test-single-digit-2-g2}, in which we can observe how a bad first decision essentially ruined the network for good. The jagged edges of the red curve in Figure \ref{fig:mnist-test-single-digit-2} correspond with the positive and negative slopes of the cluster of bad pruning decisions in \ref{fig:mnist-test-single-digit-2-g2}. Once again, these are not necessarily bad decisions, but the starting point is already bad and this cannot be recovered without re-training the network.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-2-g1.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the first-order iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-2-g1}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{png/mnist-test-single-digit-2-g2.pdf}
\caption{Error surface of the network output in log space (left) and real space (right) with respect to each candidate neuron chosen for removal using the second-order iterative re-ranking removal criterion}
\label{fig:mnist-test-single-digit-2-g2}
\end{figure}

\subsubsection{Aside: Implications of This Experiment}

From the three examples above, we see that in each case, starting from a sub-optimal network, a brute force removal technique consistently improves performance for the first few pruning iterations, and the sum of squared errors does not degrade beyond the starting point until around 60-80\% of the neurons have been removed. This is only possible if we have an essentially strict dichotomy between the roles of different neurons during training. If the network needs only 20-40\% of the neurons it began with, the training process is essentially dominated by the task of canceling the residual noise of redundant neurons. Furthermore, the network can get stuck in training with redundant units and distort the final output. This is strong evidence of our thesis that the learning representation is neither equitably nor evenly distributed and that most of the neurons which do not directly participate in the learning representation can be removed without any retraining.

@@

As can be seen from the experiments on MNIST, even though the 2nd-order approximation criterion is consistently better than 1st-order, its performance is not nearly as good as brute force based ranking, especially beyond the first layer. What is interesting to note is that from some other experiments conducted on toy datasets (predicting whether a given point would lie inside a given shape on the Cartesian plane), the performance of the 2nd-order method was found to be exceptionally good and produced results very close to the brute force method. The 1st-order method, as expected, performed poorly here as well. Some of these results are illustrated in Figure \ref{fig:diamond}.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.49\linewidth]{png/diamond-iterative-rerank-method.pdf}
\includegraphics[width=0.49\linewidth]{png/diamond-iterative-rerank-accuracy.pdf}
\includegraphics[width=0.49\linewidth]{png/rshape-iterative-rerank-method.pdf}
\includegraphics[width=0.49\linewidth]{png/rshape-iterative-rerank-accuracy.pdf}
\caption{Degradation in squared error (left) and classification accuracy (right) after pruning a 2-layer network using the iterative re-ranking algorithm on a toy ``diamond'' shape dataset (top) and a toy ``random shape'' dataset (below); (\textbf{Network:} 2 layers, 50 neurons/layer, 10 outputs, logistic sigmoid activation, starting test accuracy: 0.992(diamond); 0.986(random shape)}
\label{fig:diamond}
\end{figure}

%This huge variation in results from MNIST might be attributed to the fact that these toy datasets had only 2 output classes (as opposed to 10 classes in MNIST), but it certainly warrants further investigation.
